A Switching Planner for Combined Task and Observation Planning

### 2010-11-21 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
I am burnt out. Perhaps I will get some more energy later tonight.
Moritz, I think I am representing your work correctly, but part of the pain is mulling over your Python, and staring at a cloud...

 - Charles

Any POMDP with stochastic actions can be compiled into an equivalent
deterministic-action POMDP with all the action uncertainty expressed
in the starting-state distribution~\cite{ng:Jordan:2000}. In our
setting, that of finite-horizon planning, this compilation yields a
finite flat representation of the POMDP. Therefore, although our
current implementation does not explicitly support it, our approach
generalises to POMDPs with stochastic action models.

@@

\noindent The interpretation of such an expression can be given
according to a recursive notion of {\em visitation} over terms. Here,
an atom is {\em visited} if its conjunctive parent is visited, and a
conjunctive term is visited if all its immediate subterms are
visited. A probabilistic term is visited if exactly one of its
subterms, $T_i$, is visited. Each valid visitation of the root term
according to this recursive definition encapsulates a starting state,
along with the probability it occurs. The former corresponds to the
union of all visited atoms, and the latter corresponds to the product
of $p_i$ entries on the visited subterms of probabilistic
elements. Making these ideas concrete, our example yields the
following flat state distribution:

@@

We now describe a continual planning approach that switches between
sequential and decision-theoretic planning. The problem at hand is a
POMDP described in DTPDDL.

@@

Our approach uses sequential planning to compute a linear plan, and
one trace of that plan's execution. The latter encapsulates
assumptions about: (1) the true underlying state, (2) how the plan's
execution will progress, and (3) the possibility of eventually holding
strong beliefs about the truth values of specific state
propositions. Sequential planning proceeds in a continual manner until
the verification of an assumption about a runtime state proposition is
scheduled. At that point the system switches to a decision-theoretic
planner that is used to tailor sensory processing by the robot to test
the validity of the assumptions so far.

The first actions sequenced by the planner are virtual, and correspond
to identifying an abstract start state. That prefix corresponds to a
{\em relaxed visitation} of the root term of the DTPDDL $(:init)$
block. In the relaxed case, we can modify the initial definition so
that a conjunctive term is visited provided all its atomic subterms
are visited, and zero or more of its probabilistic terms are
visited. Writing \#\ if the value of a proposition is unspecified by a
the {\em relaxed visitation}, the $\pp{R2D2}$ example has the
following abstract starting states:

\small
\begin{tabular}{cccc}
\hline
Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
\hline
.24 & kitchen & office & office \\
.3 & kitchen & \# & office \\
.8 & kitchen & office & \# \\
.2 & kitchen & kitchen & \# \\
.7 & kitchen & \# &  kitchen\\
.06 & kitchen & kitchen & office \\
.56 & kitchen & office & kitchen \\
.14 & kitchen & kitchen & kitchen \\
\hline
\end{tabular}
\normalsize

>>>>

@@

\noindent The interpretation of such an expression can be given
according to a recursive notion of {\em visitation} over terms. Here,
an atom is {\em visited} iff its conjunctive parent is visited, and a
conjunctive term is visited iff all its immediate subterms are
visited. A probabilistic term is visited iff its conjunctive parent is
visited, and exactly one of its subterms, $T_i$, is visited. Each
valid visitation of the root term according to this recursive
definition encapsulates a starting state, along with the probability
it occurs. The former corresponds to the union of all visited atoms,
and the latter corresponds to the product of $p_i$ entries on the
visited subterms of probabilistic elements. Making these ideas
concrete, our example yields the following flat state distribution:

@@

We now describe a continual planning approach that switches between
sequential and decision-theoretic planning.

In this work, the problem at hand is a POMDP described in DTPDDL. From
hereon, we restrict our attention to POMDPs with deterministic
actions. Therefore it is worthwhile noting that any POMDP with
stochastic actions can be compiled into an equivalent
deterministic-action POMDP, where all the original action uncertainty
is expressed in the starting-state
distribution~\cite{ng:Jordan:2000}. In our setting, that of
finite-horizon planning, such a compilation yields a finite flat
representation of the original POMDP, only with deterministic
actions.

PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
HOW WE DO BELIEF REVISION ACROSS SWITCHING.

\Omit{
  Therefore, although our current implementation does not
  explicitly support it, our approach generalises to POMDPs with
  stochastic action models.
}

@@

Our approach uses sequential planning to compute a linear plan, and
one trace of that plan's execution. The latter encapsulates
assumptions about: (1) the true underlying state, (2) how the plan's
execution will progress, and (3) the possibility of the agent
eventually holding strong beliefs about the truth values of specific
state propositions. Sequential planning proceeds in a continual manner
until the verification of an assumption about a runtime state
proposition is scheduled. At that point the system switches to a
decision-theoretic planner that is used to tailor sensory processing
by the robot to test the validity of the assumptions made so far.

In a deterministic-action POMDP, all the uncertainty in state and
action is expressed in the $(\pp{:init})$ declaration. Our approach
uses the structure of that, as it occurs in the problem description,
do define a set of abstract state-assumptions available to 
sequential planning.  
%%
Writing \#\ if the value of a proposition is unspecified in the
assumption, for the $\pp{R2D2}$ example we have that the following
assumptions are available:

\small
\begin{tabular}{cccc}
\hline
Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
\hline
.24 & kitchen & office & office \\
.06 & kitchen & kitchen & office \\
.56 & kitchen & office & kitchen \\
.14 & kitchen & kitchen & kitchen \\
.7 & kitchen & \# &  kitchen\\
.3 & kitchen & \# & office \\
.8 & kitchen & office & \# \\
.2 & kitchen & kitchen & \# \\
1.0 & kitchen & \# & \# \\
\hline
\end{tabular}
\normalsize

\noindent A valid assumption in our setting corresponds to a {\em
relaxed} visitation of the root term of $(:init)$. In this relaxed
case, a conjunctive term is visited iff its atomic subterms are
visited, and zero or more of its immediate probabilistic subterms are
visited. The starting state, $\state_0$, for sequential planning is
the abstract state with probability $1.0$ -- E.G., for $\pp{R2D2}$
$s_0 \equiv \{(= (is-in R2D2) kitchen)\}$.

In order to make assumptions available to the sequential planner, we
add one action $\actions(p_i;(T_i))$ to the problem for each element,
$p_i (T_i)$, of each probabilistic term from $(:init)$. The physics
of these actions is as follows. $\actions(p_i;(T_i))$ can be executed
if no $\actions(p_j;(T_j))$, $j \neq i$, has been executed from the
same probabilistic term, and, either
$(\pp{probabilistic}~..p_i~(T_i)..)$ is in the root conjunct, or it
occurs in $T_k$ for some executed $\actions(p_k;(T_k))$.
%%
Executing $\actions(p_i;(T_i))$ in $s$ effects a transition to a
successor state $s^v$ with probability $p_i$, and $s^\bot$ with
probability $1 - p_i$. Here, $s^v$ is characterised by the union of
$s$ and atomic terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
$\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
$\transProb(\state^\bot, \action, \state^\bot) = 1$.

Because $\state_0$ is abstract, and because actions can have
preconditions that are not necessarily true in the belief-state, we
must make two further adjustments to our sequential continual
planning. First, if $\action\in\actions$ has a precondition that is
not true with probability $1$ in the current state, then before it is
executed blindly, continual planning switches to a decision-theoretic
mode, that shall be discussed in detail below. Therefore, by
scheduling $\action$ the sequential planner makes a naive assumption
about the possibility of the agent eventually holding strong beliefs
about the truth values of specific state propositions. Second, if the
truth value of $\prop$ is not known, and no assumption is made about
it in the current abstract state, then if $\prop \in \poss(\action)$
we have that action $\action$ is not applicable.

Finally, we come to the issue of optimisation criteria. Here, we
propose the value of an $N$ horizon trace ---something we expect the
sequential planner to maximise--- corresponds to the maximal value of
that trace. In detail, where $p_i$ is the probability that the
$i^{th}$ execution is not a transition to $s^\bot$, we value a trace
according to:

\begin{equation}\label{eq:tracevalue}
V(s_0, a_0, s_1, a_1,.., s_N) =  \prod_{i=1..N} p_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
\end{equation}

\noindent In goal directed problems, where the only non-zero reward is
received on the first transition to a goal state,
Eq~\ref{eq:tracevalue} corresponds exactly to the criteria of the
infamous degenerate hindsight system
\system{FFR$_a$}~\cite{yoon:etal:2007}. Moreover, form the perspective
of gradient-based reinforcement learning systems, such as William's
REINFORCE and GPOMDP, our optimal trace yields a maximal gradient
contribution under reasonable assumptions.
